---
title: "p8105_hw3_yh3554"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
```


# Problem 1

Load the data from `p8105.datasets`.

```{r}
library(p8105.datasets)
data("instacart")
```
#### Exploring the dataset

```{r}
head(instacart %>% 
  group_by(department, aisle, product_name) %>%
  summarize(n = n()) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(desc(n)))
```

#### Summary of instacart dataset
The instacart dataset is online grocery data from the Instacart in 2017. It has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each observation from the dataset is a product from an order. The dataset includes identifier, product information, order number, order from cart or not, reorder from past or not, customer identifier, time variables the ordered was placed, aisle information, and department information.\
The key variables are user identifier, product identifier, product name, aside, department name, day which ordered in a week, hour which ordered in a day, day since prior order, and order numbers. There are `r length(unique(instacart$user_id))` customers using the app. The mean hour which ordered in a day is `r round(mean(instacart$order_hour_of_day))`, the earliest hour is `r min(instacart$order_hour_of_day)`, the lastest hour is `r max(instacart$order_hour_of_day)`. The most popular products are banana, bag of organic bananas, organic, strawberries, raby spinach, raspberries, and bluebeeries which are all from produce department.\
For example, Row 1 indicates one customer purchased `r instacart$product_name[1]` at `r instacart$order_hour_of_day[1]` o'clock on Thursday. This customer had made `r instacart$order_number[1]-1` purchases before this order. This customer picked the product from `r instacart$aisle[1]` aisle in the `r instacart$department[1]` department. 

#### Total number of aisles and most popular aisles.

Apply `distinct()` to find the numbers of aisles, and use `count()` to find the most popular aisles.
```{r}
p1_df1 <- instacart %>% 
  select(aisle, department, aisle_id) %>%
  arrange(aisle) %>%
  distinct(aisle, .keep_all = TRUE)

p1_df2 <- instacart %>%
  count(aisle, name = "n_obs") %>%
  arrange(desc(n_obs))
head(p1_df2)
```

There are `r nrow(p1_df1)` aisles, and the aisles have the most ordered item are `r p1_df2$aisle[1]`, `r p1_df2$aisle[2]`, and `r p1_df2$aisle[3]` which corresponding to `r p1_df2$n_obs[1]`, `r p1_df2$n_obs[2]`, and `r p1_df2$n_obs[3]` items.

#### Plot of number of items ordered for each aisles with more than 10,000 items

```{r}
p1_df3 <- instacart %>%
  count(aisle, name = "n_items") %>%
  arrange(desc(n_items)) %>%
  filter(n_items >= 10000)

numbers_items_barplot <- p1_df3 %>%
  ggplot(aes(reorder(aisle, n_items), n_items, fill = aisle)) + 
  geom_bar(stat = "identity") + 
  labs(
    x = "aisle",
    y = "number of items",
    title = "Boxplot of number of items in each aisle"
    ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        legend.position = "none")

numbers_items_barplot
```

#### Table of three most popular items in "baking ingredients", "dog food care", and "packaged vegatables fruits".

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarize(n = n()) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable(digit = 0, caption = "most popular items in each aisles")
```

#### Table of mean hour of the day at Pink Lady Apples and Coffee Ice Cream are ordered in a week

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  mutate(order_dow = recode(order_dow, "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", 
                   "4" = "Thursday", "5" = "Friday", "6" = "Saturday")) %>%
  pivot_wider(names_from = "order_dow", values_from = "mean_hour") %>%
  knitr::kable(digits = 2, caption = "mean hour of given product are ordered in a week")
```


# Problem 2

Load, tidy, and wrangle the accel data. 

```{r}
p2_df1 <- read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    weekday_weekend = recode(day, "Sunday" = 0, "Saturday" = 0, "Monday" = 1, "Tuesday" = 1, "Wednesday" = 1, "Thursday" = 1, "Friday" = 1)) %>%
  select(week, day_id, day, weekday_weekend, everything()) %>%
  rename_all(~stringr::str_replace(.,"activity_","")) %>%
  pivot_longer("1":"1440", names_to = "n_min", values_to = "n_activity") %>%
  mutate(n_min = as.numeric(n_min))
  
head(p2_df1)
```

The p2_df1 dataset has `r nrow(p2_df1)` rows and `r ncol(p2_df1)` variables. The variables are week, day id, day, weekday or weekend (weekday is 1, weekend is 0), minute number, and number of activity.

#### Table of total activity in a day

Generate a new data from p2_df1 the tidied dataset that agregate the number of activity in a day.

```{r}
p2_df2 <- p2_df1 %>% 
  group_by(week, day_id, day) %>%
  summarise(tot_activity = sum(n_activity))

table_1 <- p2_df2 %>%
   knitr::kable()

table_1
```

Create a trend plot to see if there is any pattern.

```{r}
p2_df2$day <- factor(p2_df2$day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) 

p2_df2 %>% ggplot(aes(x = day, y = tot_activity, color = week, group = week)) +
  geom_point() + geom_line() + facet_grid(~week) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
        legend.position = "none")
```

Based on the trend plot, we can see the patient reached to peak on weekend (either Saturday or Sunday) in the first two weeks, and the patient was more active on weekday in the last three weeks.



